NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-02-22 Manuscript/manuscriptPS.tex
Updates to text

The headgroup order parameters of DOPS and POPS bilayers 
from different simulation models are compared with the experimental
data in Fig. \ref{HGorderParametersPS}. In line with the previous
study for PC lipids \cite{botan15}, the glycerol backbone order 
parameters in CHARMM36 roughly agree with the experimental data,
while significant discrepancies are observed in other models at
least for one glycerol backbone carbons.
\todo{Discussion to be finished once all the results are in the plot.}

The glycerol backbone structures from CHARMM36 and
Slipids simulations visualized in Fig. \ref{glycerol_buslaev} reveal the
missing structures in Slipids model, which probably lead to the incorrect
order parameters. For the headgroup $\alpha$ and $\beta$ carbon order parameters of PS lipids,
the tested models perform less well than for PC headgroup in previous
study \cite{botan15}. None of the tested models give a satisfactory 
agreement with experiments for the headgroup order paramaters of 
PS lipids. Based on the subjective ranking shown in Fig. \ref{comparisonTablePS},
the CHARMM36 is the best performing model for both lipids.
However, the total deviation from the experiments for the PS lipids is significantly
higher for PS lipids (8) than for PC (3) \cite{botan15}. Therefore the 
interpretation of structural differences between PC and PS lipids 
from current MD simulation models is very challenging.

\todo{Dihedral angle distribtions in Fig. \ref{dihedrals} should be analyzed from different models 
  and included in the discussion. Maybe also figures similar to \ref{glycerol_buslaev}.}

@@

In contrast to PC lipids discussed previously \cite{botan15,catte16},
systems with negatively charged lipids always contains counterions.
Since the ion binding changes the headgroup order parameters \cite{seelig87,catte16},
the counterion distribution have to be taken into account also
when additional ions are not present.
The Na+ density profiles from different simulation are shown in Fig. \ref{NAdensPOPS}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/NAdensPOPS.eps}
  \caption{\label{NAdensPOPS}
    Sodium density of POPS lipid bilayer along the membrane normal from
    simulations with different force fields.
  }
\end{figure}

>>>>

The headgroup order parameters of DOPS and POPS bilayers 
from different simulation models are compared with the experimental
data in Fig. \ref{HGorderParametersPS}. In line with the previous
study for PC lipids \cite{botan15}, the glycerol backbone order 
parameters in CHARMM36 roughly agree with the experimental data,
while significant discrepancies for glycerol backbone carbons
are observed in other models.
\todo{Discussion to be finished once all the results are in the plot.}

Based on the subjective ranking shown in Fig. \ref{comparisonTablePS},
the CHARMM36 is the best performing model for both lipids.
However, none of the tested models give a satisfactory 
agreement with experiments for the headgroup order paramaters of 
PS lipids. The total deviation from the experiments ($\Sigma$ in Fig. \ref{comparisonTablePS})
for the best performing CHARMM36 model is larger for PS lipids (8)
than for PC (3) \cite{botan15}. Therefore the 
interpretation of structural details of PS headgroup or
differences between PC and PS lipids is very challenging 
from the current MD simulation models.
Figures \ref{dihedralsGLY} and \ref{dihedralsHG} show dihedral
angle distributions calculated from different models. 
The glycerol backbone structures from CHARMM36 and
Slipids simulations visualized in Fig. \ref{glycerol_buslaev} reveal the
missing structures in Slipids model, which probably lead to the incorrect
order parameters. For the headgroup $\alpha$ and $\beta$ carbon order parameters of PS lipids,
the tested models perform less well than for PC headgroup in previous
study \cite{botan15}.
\todo{Discussion will be finished when we have all the data in
  Figs. \ref{HGorderParametersPS}, \ref{dihedralsGLY} and \ref{dihedralsHG}}

@@

Since systems with negatively charged PS lipids always contain counterions
and the ion binding presumably affects the PS headgroup
order parameters, the discussion of ion binding affinity and
headgroup structure cannot be separated as done previously
for PC lipids \cite{botan15,catte16}. The density profiles
of counterions along membrane normal from different simulations
are shown in Fig. \ref{NAdensPOPS}.
Significant differences are observed in counterion
densities between different simulation models. 
\todo{We should try to figure out which one of these are more
  realistic. One option could be to use PS headgroup order parameter
  data measured as a function of NaCl \cite{roux86}. Another
  approach could be to use the area per molecule, which is suggested to
  depend on the binding affinity of counterions \cite{pandit02,mukhopadhyay04,pedersen06}. The
  experimental value for area per molecule is available at \cite{pan14}.}
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/NAdensPOPS.eps}
  \caption{\label{NAdensPOPS}
    Counterion densities of POPS lipid bilayer along the membrane normal from
    simulations with different force fields.
  }
\end{figure}
